---
title: "exp2_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library("ggplot2")
library("dplyr")
library("patchwork")
library("brms")
library("lme4")
library("tidyverse")
library("tidybayes")
library("scales")
library("faintr")
library("ggeffects")
library("ggh4x")
```

# Preprocessing

## minimal possible RTs for data exclusion (seperate if and because)

## text-based experiment
```{r}
data_text <- read.csv(file="../data/exp2_prolific_all.csv",header=TRUE,sep=",")
data_text <- data_text[data_text$experiment=='text',c('submission_id','condition','structure','response','expect_ans','responseTime')]
data_text<-data_text[!(data_text$submission_id==3819),]

data_text <- data_text %>% mutate(submission_id = factor(submission_id),condition = factor(condition),structure = factor(structure))
```
```{r}
data_text %>% filter(!submission_id %in% final_out$submission_id)
```
## text-based experiment bar plot
```{r}
 
df.plot = data_text %>%  
  filter(!submission_id %in% final_out$submission_id) %>% 
  mutate(normal = ifelse(structure == "disjunctive",response == expect_ans,response != expect_ans)) %>% mutate(normal = ifelse(normal == TRUE, 1, 0))

text_plot = ggplot(data = df.plot,
       mapping = aes(x = condition, 
                     y = normal,fill=structure,
                     group = structure
                     )) +
  geom_hline(yintercept = 0.5, 
             linetype = 2) + 
  stat_summary(fun.y = "mean",
               geom="bar",
               position = position_dodge2(width = 0.9),
               size=0.1
               #size = 1.5,shape=21
               ) +
  stat_summary(fun.data = "mean_cl_boot",
               geom = "errorbar",
               #geom="bar",
               position = position_dodge(width=0.9),
               width=0.2) +
  labs(x = "type of condition",
       y = "% normal cause selections" #, title = "Participants' selections"
       ) +
  scale_fill_brewer(palette = "Set3") +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = str_c(seq(0, 100, 25), "%"),
                     expand = expansion(mult = c(0.05, 0))) +
  coord_cartesian(clip = "off",
                  ylim = c(0, 1)) + 
  theme(text = element_text(size = 15),
        legend.position = "bottom",
        #panel.grid = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_blank(),
        #plot.margin = margin(t = 1, l = 0.2, r = 0.2, unit = "cm"),
        plot.title = element_text(vjust = 4,
                                  color = "gray40",
                                  hjust = 0.5)) +
  theme(axis.line = element_line(color="black", size = 0.6))


text_plot
```

## dataset exclude participants with technical issues
```{r}
data_all <- read.csv(file="../data/exp2_prolific_all.csv", header=TRUE,sep=",")
data_all <- data_all[data_all$experiment=='main',c('submission_id','condition','pA','pB','pos','structure','response','expect_ans','responseTime')]
data_all<-data_all[!(data_all$submission_id==3819),]

data_all <- data_all %>% mutate(submission_id = factor(submission_id),condition = factor(condition),structure = factor(structure))

```
```{r}
hist(data_all[data_all$condition=="if","responseTime"], breaks = 100)

```

```{r}

mu <- data_all %>% 
  group_by(condition) %>% 
  dplyr::summarize(Mean=mean(responseTime))

ggplot(data_all, aes(x=responseTime/1000, fill=condition, color=condition)) +
  geom_histogram(position="identity", alpha=0.3) +
  scale_x_continuous(labels = label_number(),breaks=pretty_breaks(n=50)) +
  scale_y_continuous(breaks=pretty_breaks(n=30)) +
  geom_vline(data=mu, aes(xintercept=Mean/1000, color=condition),linetype="dashed",show.legend = FALSE) +
  geom_text(data=mu,aes(x=Mean/1000,y=c(300,280),label=round(Mean/1000,digits=3)),show.legend = FALSE) +
  labs(y="count", title = "Distribution of RTs (all data)") +
  theme(axis.text.x = element_text(angle = 90,vjust=1))
```

## exclude participants with minimal possible RTs (arbitary threshold)
```{r}
out_ind_if <- which(data_all$condition=="if" & data_all$responseTime < 2700)
out_ind_bc <- which(data_all$condition=="because" & data_all$responseTime < 5000)

rt_out <- data_all[union(out_ind_if,out_ind_bc),]
rt_out[,c('submission_id','condition','structure','responseTime')] 
#rt_out %>% filter(expect_ans != response)
```
```{r}
final_out <- rt_out %>%
  count(submission_id) %>%
  filter(n>=2)
final_out # 
```
# Visualization
```{r}
cond_disj_conj <- data_all %>% filter(condition=='if' & (structure=='disjunctive' | structure=='conjunctive'))  

filler_disj_conj <- data_all %>% filter(condition=='because' & (structure=='disjunctive' | structure=='conjunctive')) 
```

## Percentage of match with expectations 

| chain | common_both | common_single | conjunctive | disjunctive | single |
|:---:|:---:|:---:|:---:|:---:|:---:|
| the ball in the middle|the initially moving ball | the initially moving ball | the ball with a low possibility to pass | the ball with a high possibility to pass|the ball in the same row with E |

The max for each column is 2 x num_participants
```{r}
total_count <- data_all %>% 
  filter(!submission_id %in% final_out$submission_id) %>%
  count(condition, structure, expect=(response==expect_ans)) 
total_count
```
```{r}
data_plot = data_all %>%  
  filter(!submission_id %in% final_out$submission_id) %>% 
  mutate(expect = ifelse((response==expect_ans), 1, 0))

ggplot(data=data_plot,aes(x = structure, 
                     y = mean(expect), fill=condition,
                     group = condition)) +
  labs(y="match%", title = "Percentage of match with expectations") +
  geom_bar(position="dodge", stat="identity") 
  
```
```{r,out.height=100%,out.width=100%}
total_count %>%
  filter(expect == TRUE) %>%
  ggplot(aes(x=structure, y=n/90, group=condition, fill=condition)) +
  labs(y="match%", title = "Percentage of match with expectations") +
  geom_bar(position="dodge2", stat="identity") 
  
  #geom_text(aes(label=round(n/90, digits = 3)),vjust=0,position = position_dodge(.9))
```

## individual 
```{r}
individual_count <- data_all %>%     
  #filter(!submission_id %in% final_out$submission_id) %>%
  count(submission_id, condition, structure) 
individual_count
```
```{r}
individual_count <- data_all %>%     
  #filter(!submission_id %in% final_out$submission_id) %>%
  count(submission_id, condition, structure, expect=(response==expect_ans)) 

individual_count
```
```{r, fig.height = 5, fig.width = 10}
individual_count %>%
  #filter(expect == TRUE) %>%
  ggplot(aes(x=structure, y=n, fill=expect)) + 
  geom_bar(position="stacked") +
  #facet_grid(~condition) +
  #geom_text(aes(label=n),vjust=0,position = position_dodge(.9)) +
  facet_nested(~submission_id+condition)
  #facet_wrap(~submission_id,ncol=5) +
  theme(axis.text.x = element_text(angle = 20)) 
```

## count normal event in disjunctive and conjunctive conditions

```{r}
count_disj_conj <- data_all %>% 
  filter(structure=='disjunctive' | structure=='conjunctive') %>% 
  filter(!submission_id %in% final_out$submission_id) %>%
  count(condition, structure, normal = ifelse(structure == "disjunctive",response == expect_ans,response != expect_ans))

count_disj_conj
```
```{r}
count_disj_conj %>%
  filter(normal==TRUE) %>%
  ggplot(aes(x=structure, y=n/90, group=condition,fill=condition)) +
  labs(y="normal%", title = "Percentage of normal choice") +
  geom_bar(position="dodge", stat="identity") +
  #stat_summary(fun.data = mean_cl_boot, geom = "errorbar",width = 0.2, size = 0.8, position = position_dodge(0.8)) +
  geom_text(aes(label=round(n/90, digits = 3)),vjust=0,position = position_dodge(.9))

```

# brms after data exclusion


```{r}
data_disj_conj <- data_all %>% filter(structure=='disjunctive' | structure=='conjunctive') %>% filter(!submission_id %in% final_out$submission_id)
```


```{r}
df.plot = data_disj_conj %>% mutate(normal = ifelse(structure == "disjunctive",response == expect_ans,response != expect_ans)) %>% mutate(normal = ifelse(normal == TRUE, 1, 0))

exp2_plot = ggplot(data = df.plot,
       mapping = aes(x = condition, 
                     y = normal,fill=structure,
                     group = structure
                     )) +
  geom_hline(yintercept = 0.5, 
             linetype = 2) + 
  stat_summary(fun.y = "mean",
               #geom = "pointrange",
               geom="bar",
               position = position_dodge2(width = 0.9),
               size=0.1
               #size = 1.5,shape=21
               ) +
  stat_summary(fun.data = "mean_cl_boot",
               geom = "errorbar",
               #geom="bar",
               position = position_dodge(width=0.9),
               width=0.2) +
  labs(x = "type of condition",
       y = "% normal cause selections" #, title = "Participants' selections"
       ) +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = str_c(seq(0, 100, 25), "%"),
                     expand = expansion(mult = c(0.05, 0))) +
  coord_cartesian(clip = "off",
                  ylim = c(0, 1)) + 
  theme(text = element_text(size = 15),
        legend.position = "bottom",
        #panel.grid = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_blank(),
        #plot.margin = margin(t = 1, l = 0.2, r = 0.2, unit = "cm"),
        plot.title = element_text(vjust = 4,
                                  color = "gray40",
                                  hjust = 0.5)) +
  theme(axis.line = element_line(color="black", size = 0.6))


exp2_plot
```


```{r}
data_plot = data_all %>%  
  filter(!submission_id %in% final_out$submission_id) %>% 
  mutate(expect = ifelse((response==expect_ans), 1, 0))

exp2_plot2 = ggplot(data = data_plot,
       mapping = aes(x = structure, 
                     y=expect,
                     fill=condition,
                     group = condition
                     )) +
  geom_hline(yintercept = 0.5, 
             linetype = 2) + 
  #geom_bar(stat = "identity", position = position_dodge(width=0.3))+
  stat_summary(fun.y = "mean",
               geom="bar",position=position_dodge2(width=0.9)) +
  stat_summary(fun.data = "mean_cl_boot",
               #geom = "pointrange",
               geom="errorbar",
               position = position_dodge(width=0.9),
               #size = 1.5,
               width=0.2
               ) +
  labs(x = "type of structure",
       y = "% match with expectations"#,title = "Participants' selections"
       ) +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = str_c(seq(0, 100, 25), "%"),
                     expand = expansion(mult = c(0.05, 0))) +
  coord_cartesian(clip = "off",
                  ylim = c(0, 1)) + 
  theme(text = element_text(size = 15),
        legend.position = "bottom",
        #panel.grid = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 12),
        plot.margin = margin(t = 1, l = 0.2, r = 0.2, unit = "cm"),
        plot.title = element_text(vjust = 4,
                                  color = "gray20",
                                  hjust = 0.5)) +
  theme(axis.line = element_line(color="black", size = 0.6))

exp2_plot2

```
```{r}
data_all_text <- read.csv(file="../data/exp2_prolific_all.csv", header=TRUE,sep=",")
data_all_text <- data_all_text[data_all_text$experiment=='main' | data_all_text$experiment=='text',c('submission_id','experiment','condition','pA','pB','pos','structure','response','expect_ans','responseTime')]
data_all_text<-data_all_text[!(data_all_text$submission_id==3819),]

data_all_text <- data_all_text %>% mutate(submission_id = factor(submission_id),condition = factor(condition),structure = factor(structure),experiment=factor(experiment))

```
```{r}
data_all_disj_conj <- data_all_text %>% 
  filter(!submission_id %in% final_out$submission_id) %>%
  filter((structure=='disjunctive' | structure=='conjunctive'))

data_all_disj_conj
```
```{r}
all_brm_out <- brm(formula = normal~1+ structure*condition+structure*experiment+(1+structure+condition+experiment|submission_id), data = data_all_disj_conj %>% mutate(normal = ifelse(structure == "disjunctive",response == expect_ans,response != expect_ans)), family = bernoulli)
```
```{r}
saveRDS(all_brm_out,"./brms-model_full.rds")

```
```{r}
summary(all_brm_out)
```
```{r}
#brms::hypothesis(all_brm_out, "structuredisjunctive > 0")
#get_cell_definitions(all_brm_out)
draws_cond <- cbind(extract_cell_draws(all_brm_out, condition == 'if', colname = 'if'),
                      extract_cell_draws(all_brm_out, condition == 'because', colname = 'because'))

draws_cond %>% 
  pivot_longer(cols = everything(), names_to = 'group', values_to = 'draws') %>% 
  ggplot(aes(x = draws, color = group, fill = group)) +
  geom_density(alpha = .4)
```
```{r}
compare_groups(model=all_brm_out,lower=experiment=="main"&condition=='if' & structure=='conjunctive')
```
```{r}
compare_groups(model = all_brm_out, 
               higher = experiment=="text"&condition=='if' & structure=='disjunctive',
               lower = experiment=="main"&condition=='because' & structure=='disjunctive')
```


```{r}
compare_groups(model = all_brm_out, 
               higher = experiment=="main"&condition=='if' & structure=='disjunctive',
               lower = experiment=="main"&condition=='if' & structure=='conjunctive')
```
```{r}
compare_groups(model = all_brm_out, 
               higher = experiment=="text"&condition=='if' & structure=='disjunctive',
               lower = experiment=="main"&condition=='because' & structure=='disjunctive')
```
```{r}
#pp_check(all_brm_out)
pp_check(all_brm_out, type = "stat", stat = 'mean', ndraws = 100)
#hypothesis(all_brm_out, "structuredisjunctive > 0")
```
```{r}
all_brm_out <- readRDS("./brms-model_full.rds")
```
```{r}
mcmc_plot(all_brm_out)
```
```{r}
get_cell_definitions(all_brm_out)
```